---
title: "KRS"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{KRS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

The [Kew Reconciliation Service (KRS)](http://data1.kew.org/reconciliation/about/IpniName) allows a user submit a taxon for matching against [IPNI](https://ipni.org/). 

The reconciliation service is an Open Refine style API that allows matching of a single name against IPNI. The matching is done by a series of transformations configured to botanical names
in IPNI. These transformations are detailed [here](https://github.com/RBGKew/String-Transformers#string-transformers), I think.

It appears that KRS is the service that sits behind [KNMS](KNMS.html). KNMS allows matching of
batches of names in one request but does not allow matching to different parts of a name. If you have a set of names to match and just want to do simple matching, I'd use KNMS. But if you want
to specify which parts of the names to match on, I'd use KRS.


```{r setup}
library(kewr)
library(dplyr)
library(tidyr)
```

## Matching names

To use KRS, you can just submit a single name for matching.

```{r}
match <- match_krs("Solanum sanchez-vegae S.Knapp")
match
```

This also works without the author string included:

```{r}
match <- match_krs("Solanum sanchez-vegae")
match
```

The match results can be converted to a table for easier inspection.

```{r}
tidy(match)
```

## Matching parts of a name

The reconciliation service provides a specification for matching to different parts
of a botanical name. This is described in detail [here](http://data1.kew.org/reconciliation/about/IpniName).

For example, if we want to match to the genus name *Myrcia*, we could submit a
simple request like before.

```{r}
match <- match_krs("Myrcia")
match
```

But this has returned more than one result. We can be more specific by matching 
to the genus and the author.

```{r}
match <- match_krs(list(genus="Myrcia", author="DC"))
match
```

Which has narrowed it down more.

We can specify a match for every part of a name like this.

```{r}
match <- match_krs(list(genus="Myrcia", species="magnolifolia", infra="latifolia",
                        author="Berg"))
match
```

This match has worked even though there's a minor misspelling of the specific epithet
and the author string. Matching to the taxon name works by a set of pre-configured
string transformations that catch some common mistakes in botanical names. The
matching to author strings is also slightly fuzzy.

This matching also handles different taxonomic ranks using ordered epithets, where
the highest rank is specified as `epithet_1` down to `epithet_3`.

```{r}
match <- match_krs(list(epithet_1="Solanaceae"))
match
```

This also works for infrageneric names.

```{r}
match <- match_krs(list(epithet_1="Acacia", epithet_2="Aculeiferum", author="Vassal"))
match
```

It should be noted that these last two examples give a score lower than 100,
because they return more than one match.

## Matching more than one name

If you want to do simple matching to more than one name, it might be easier to use
KNMS.

If you want to match the individual parts of multiple names, you can apply the
matching function to the rows of a data frame, using `dplyr::rowwise`.

```{r}
names <- tibble(
  genus=c("Poa", "Myrcia", "Solanum"),
  species=c("annua", "almasensis", "sanchez-vegae"),
  author=c("L.", "NicLugh.", "S.Knapp")
)


matches <-
  names %>%
  rowwise() %>%
  mutate(match=list(match_krs(list(genus=genus, species=species, author=author)))) %>%
  mutate(match=list(tidy(match))) %>%
  unnest(cols=c(match))

matches
```

