---
title: "WCVP"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{WCVP}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, warning=FALSE, message=FALSE}
library(kewr)
library(here)
library(dplyr)
```

The [World Checklist of Vascular Plants (WCVP)](https://wcvp.science.kew.org/) is a global consensus view of all known vascular plant species. It has been compiled by staff at RBG Kew in consultation with plant group experts.

The WCVP is a taxonomic database, and as such contains information like the taxonomic status and synonymy of plant names. It can be used for a number of tasks, including searching for all records with a particular name, getting a list of all accepted species in a genus, or looking up the accepted name for a species.

## Searching WCVP for a taxon name

Searching in WCVP works by exact matching. This means that a misspelled name will not return any results.

For instance, searching for `Poa anua` gets nothing:

```{r}
results <- search_wcvp("Poa anua")
results
```

But searching for the correctly spelled name will give a result:

```{r}
results <- search_wcvp("Poa annua")
results
```

Similarly, searching with partial matching does not work:

```{r}
results <- search_wcvp("Ulex e")
results
```

But searching for a genus name will return all taxa within that genus as well:

```{r}
results <- search_wcvp("Ulex")
results
```

With this search result, we can see that there are 92 records for `Ulex` but the API has only returned the first 50.

To get all results for this search, we can either do a second search to download the second page of results:

```{r}
ulex_page1 <- search_wcvp("Ulex")
ulex_page2 <- search_wcvp("Ulex", page=2)
ulex_page2
```

Or increase the maximum number of results to return:

```{r}
ulex_all <- search_wcvp("Ulex", limit=100)
ulex_all
```

In both cases, we can format the results into a tibble to make subsequent analysis easier:

```{r}
format(ulex_all)
```

```{r}
bind_rows(
  format(ulex_page1),
  format(ulex_page2)
)
```

## Getting a list of accepted names in a genus

The search function also accepts a number of keywords and filters that can be used to narrow down the results returned. A full list of keywords and filters can be found in the function documentation.

An example use of these filters is retrieving a checklist of accepted species in a genus. In the search below, the `genus` keyword to retrieve all records in the genus *Myrcia*, while the filters `accepted` and `species` narrow these records down to just accepted species:

```{r}
checklist <- search_wcvp(list(genus="Myrcia"), 
                         filters=c("accepted", "species"),
                         limit=1000)

checklist
```

## Looking up accepted names

As well as searching by name, the WCVP API can be used to look up taxonomic information for a known IPNI ID. The returned information can then be used to find all synonyms for a name, find the accepted name for a taxon, or find the parent/child taxa.

The IPNI ID can be found a few different ways, for example using the `KNMS` API to match a name to an IPNI ID. In this case we will find it using WCVP.

```{r}
search_result <- search_wcvp("Calyptranthes acevedoi", 
                             filters=c("species"))

ipni_id <- search_result$results[[1]]$id
ipni_id
```

With this, we can look the record up.

```{r}
taxon_info <- lookup_wcvp(ipni_id)
taxon_info
```
From this record, we can see that the name we had is a Homotypic Synonym. This record also contains the taxonomic information for the current accepted name:

```{r}
format(taxon_info, field="accepted")
```

## Downloading the whole WCVP

There are cases where it might be easier to download all of the WCVP, rather than make lots of requests to it. The `download_wcvp` function lets you download the whole WCVP to whatever directory you want.

```{r, eval=FALSE}
save_path <- here()
download_wcvp(save_path)
```
Previous versions of the WCVP are also available, for posterity and reproducibility. These can be downloaded by specifying which version you want.

```{r, eval=FALSE}
download_wcvp(save_path, version=1)
```

