---
title: "Mapping assessed species"
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This is a demonstration of one use for some of the kewr resources, in calculating the proportion of species that have been assessed in a country.

To do this, we'll follow these steps:

1. Retrieve a list of assessed species in one country from the IUCN Red List of Threatened Species.
2. Match names to an IPNI ID using the Kew Names Matching Service.
3. Resolve assessments to accepted names using the World Checklist of Vascular Plants.
4. Get a list of all accepted vascular plant species in the country from Plants of the World Online.
5. Calculate the proportion of assessed and threatened species in our country of interest.

## Setup 

In addition to *kewr*, we'll load:

* *dplyr* to manipulate the data
* *tidyr* to reshape data frames
* *stringr* to extract some data from strings
* *ggplot2* to make some plots of our final data

```{r setup}
library(kewr)
library(dplyr)
library(tidyr)
library(stringr)
library(ggplot2)
```

## 1. Getting assessment information

The first step in this problem is to get a list of assessed species for a country.

The most authoritative list of global species assessments is the IUCN Red List of Threatened Species. For this package, I used the `rl_sp_country` function from *rredlist* to request all assessments for taxa found in Denmark from the IUCN Red List API. I then used the `rl_search` function to request full assessment information for each taxon.

You can do the same by [installing the *rredlist* package](https://github.com/ropensci/rredlist). You'll need an API key to use the IUCN Red List API, which you can [register for here](https://apiv3.iucnredlist.org/api/v3/token). Alternatively, you can download the information you need directly from [the IUCN Red List website](https://www.iucnredlist.org/).

I've bundled up the Danish plant assessments in this package to make things easier for this analysis.

```{r view-danish-plants}
head(danish_plants)
```

In total, there are assessments for `r format(nrow(danish_plants), big.mark=",")` vascular plants from Denmark.

## 2. Match names to an IPNI ID

To match these names to an IPNI ID, we'll use the Kew Names Matching Service. The first thing we'll do, is join the taxonomic authority to the scientific name, to hopefully reduce the number of matches for each name.

```{r make-full-names}
danish_plants <-
  danish_plants %>%
  unite("full_name", scientific_name, authority, sep=" ", remove=FALSE)
```

Then we put the full names through KNMS.

```{r match-full-names}
full_matches <- match_knms(danish_plants$full_name)
full_matches
```

Most of our names had matches! Which makes things simpler. But we also got some names that returned multiple matches.

First we'll try putting just the scientific names of our missing matches through KNMS again.

```{r match-partial-names}
full_matches <- format(full_matches)

unmatched <- filter(full_matches, !matched)

to_match <- 
  danish_plants %>%
  filter(full_name %in% unmatched$submitted)

part_matches <- match_knms(to_match$scientific_name)
part_matches
```
Almost everything returned a match. We could attempt to manually match the final name, but as it's just one we'll leave it out.

Now we can join all of our matches together, link them to the IUCN taxon ID, and resolve any synonyms.

```{r resolve-matches}
part_matches <- format(part_matches)

full_matches <- 
  full_matches %>%
  filter(matched) %>%
  left_join(
    danish_plants %>% select(taxonid, full_name, category),
    by=c("submitted"="full_name")
  )
  
part_matches <- 
  part_matches %>%
  filter(matched) %>%
  left_join(
    danish_plants %>% select(taxonid, scientific_name, category),
    by=c("submitted"="scientific_name")
  )

matched_names <- bind_rows(full_matches, part_matches)

head(matched_names)
```

## 3. Resolve assessments to accepted names

Now that we have an IPNI ID attached to each assessment, we can look up the record for the taxa in WCVP. This will let us find out the taxonomic status of each name - the first step in resolving any synonymy issues.

```{r resolve-synonyms}
# wrap up the lookup_wcvp function to add a wait
f <- function(taxonid) {
  Sys.sleep(0.1)
  list(lookup_wcvp(taxonid))
}

resolved_names <- 
  matched_names %>%
  nest_by(taxonid, ipni_id) %>%
  mutate(wcvp_record=f(ipni_id))

resolved_names <-
  resolved_names %>%
  mutate(status=wcvp_record$status)
```

With the taxonomic status, we'll first remove any taxa that are unplaced. We'll also remove any non-homotypic synonyms - even if we resolve these to accepted species, we can't be sure that the assessment would be valid for the new concept, so that accepted species would not be assessed.

```{r remove-non-homotypic}
resolved_names <-
  resolved_names %>%
  filter(status %in% c("accepted", "homotypic synonym"))
```

Next we'll check if there are any taxa that still have multiple matches in WCVP.

```{r check-multiple-matches}
resolved_names %>%
  ungroup() %>%
  add_count(taxonid) %>%
  summarise(multiple_matches=sum(n > 1))
```

There are not.

So the final step is to find the accepted names for all homotypic synonyms and remove anything that is a lower rank than species.

```{r find-accepted-names}
resolved_names <-
  resolved_names %>%
  mutate(accepted_id=ifelse(status != "accepted",
                              wcvp_record$accepted$id,
                              wcvp_record$id),
         accepted_name=ifelse(status != "accepted",
                              wcvp_record$accepted$name,
                              wcvp_record$name),
         accepted_author=ifelse(status != "accepted",
                              wcvp_record$accepted$author,
                              wcvp_record$authors),
         accepted_rank=ifelse(status != "accepted",
                              wcvp_record$accepted$rank,
                              wcvp_record$rank))

resolved_names <- 
  resolved_names %>%
  select(-wcvp_record) %>%
  unnest(cols=c(data)) %>%
  rename(match_id=ipni_id)

accepted_species <-
  resolved_names %>%
  filter(accepted_rank == "Species") %>%
  ungroup()

nrow(accepted_species)
```

We are now left with `r nrow(accepted_species)` accepted species from Denmark with assessments.

## 4. Get a list of all species in the country

To calculate the number of species that are assessed in Denmark, we need a checklist of all accepted species.

We can get this from Plants of the World Online.

```{r download-powo-checklist}
checklist <- search_powo(list(distribution="Denmark"),
                         filters=c("accepted", "species"),
                         limit=5000)
checklist
```

Now we have this, we just need to join our assessments to our checklist.

```{r join-assessments-checklist}
checklist <-
  format(checklist) %>%
  mutate(ipni_id=str_extract(fqId, "[0-9\\-]+")) %>%
  select(ipni_id, family, name, author) %>%
  distinct(ipni_id, .keep_all=TRUE) %>%
  left_join(
    accepted_species %>% select(accepted_id, category),
    by=c("ipni_id"="accepted_id")
  )
```

## 5. Calculating the proportion of assessed species

And now we can calculate the proportion of species assessed in Denmark!

```{r calculate-proportions}
checklist %>%
  summarise(p_assessed=mean(! is.na(category)))
```

And make a simple bar chart of the number of species in each category.

```{r plot-bars}
iucn_colours <- c("NE"="#ffffff", 
                  "DD"="#d1d1d6", 
                  "LC"="#60c659", 
                  "NT"="#cce226", 
                  "VU"="#f9e814", 
                  "EN"="#fc7f3f", 
                  "CR"="d81e05", 
                  "EW"="#542344", 
                  "EX"="#000000")
checklist %>%
  replace_na(list(category="NE")) %>%
  mutate(category=factor(category, levels=names(iucn_colours),
                         ordered=TRUE)) %>%
  ggplot(mapping=aes(y=category, fill=category,
                     colour=category == "NE")) +
  geom_bar() +
  scale_fill_manual(values=iucn_colours, drop=FALSE) +
  scale_colour_manual(values=c(`TRUE`="black", `FALSE`=NA)) +
  scale_y_discrete(drop=FALSE) +
  guides(fill=FALSE, colour=FALSE) +
  labs(x="Number of species", y="IUCN Red List category")
```


